Splints of root systems for special Lie subalgebras

### 2015-01-21 AffineLieAlgebras/Splints/special-paper.tex
Preprint published

\usepackage{amsfonts}
\usepackage{pb-diagram}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{amsmath,amssymb,amsthm}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}

@@

Note that the image $Im(\phi)$ is not required to inherit the root system
properties except the addition rules equivalent to the addition
rules in $\Delta_{1}$ (for pre-images). Two embeddings $\phi_1$ and $\phi_2$  
can splinter $\Delta$  when the latter can be presented 
as a disjoint union of images $Im(\phi_1)$ and $Im(\phi_2)$.

Why one would consider non angle-preserving maps of root systems? Additive properties of root system
determine the structure of Verma module:
    \begin{equation*}
      M^{\mu}=U(\gf)\underset{U(\bff_{+})}{\otimes} D^{\mu}(\bff_{+})=\{(E^{-\alpha_{1}})^{n_{1}}\dots (E^{-\alpha_{s}})^{n_{s}} \left|v_{\mu}\right>\}_{\alpha_{i}\in\Delta^{+}}^{n_{i}=0,1,\dots}
    \end{equation*}
Here $\bff_{+}$ is the Borel subalgebra and $D^{\mu}$ its one-dimensional representation. 
Weyl character formula expresses character of irreducible representation as a combination of
characters of Verma modules: 
\begin{equation*}
  \mathrm{ch} L^{\mu}=\frac{\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}}{\sum_{w\in W}\varepsilon(w) e^{w\rho-\rho}}=\sum_{w\in W} \varepsilon(w)\; \mathrm{ch} M^{w(\mu+\rho)-\rho}
\end{equation*}
Here the sum is over elements of Weyl group and their actions $w\triangleright \mu
=w(\mu+\rho)-\rho$ depend on angles between roots. But we can rewrite Weyl character formula
representing Weyl group elements as the products of reflections $s_{\alpha}, \alpha \in S$ in
hyperplanes orthogonal to simple roots: $w=s_{\alpha_{1}}\cdot s_{\alpha_{2}}\dots$. Reflections act
on root system by permutations, so the composition of such an action with embedding $\phi$ is easily
obtained. The orbit of Weyl groups action $w\triangleright \mu, w\in W$ can be constructed by
subtractions of roots from the highest weight $\mu$. Let's denote Dynkin labels of $\mu$ by
$(\mu_{1},\dots \mu_{r})$. Then $\mu-\mu_{1}\alpha_{1}, \mu-\mu_{2}\alpha_{2},\dots,
\mu-\mu_{r}\alpha_{r}, \alpha_{i}\in S$ are the points of the orbit that can be obtained by
elementary reflections $s_{\alpha_{i}}, \alpha_{i}\in S$. Next set of points that are obtained by
two consecutive reflections $s_{\alpha_{i}}\cdot s_{\alpha_{j}}\triangleright\mu$ are obtained as
the subtraction $\mu-\mu_{i}\alpha_{i}-\mu_{j} (s_{\alpha_{i}}\alpha_{j})$. Continuing this way we
construct the image of singular element $\Psi=\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}$
after the embedding $\phi$ \cite{2011arXiv1111.6787L}. The multiplicities in the character of Verma
module are unchanged by the embedding since they are determined by additive properties of the roots.
So we see that weight multiplicities of irreducible modules are also preserved by the embedding.

@@

Assume that root system $\Delta_{\gf}$ splinters as $\Delta_{\gf}=\Delta_{\af} \cup
\phi(\Delta_{\sfr})$, where $\phi$ is an embedding of root system $\Delta_{\sfr}$ of some semisimple
Lie algebra $\sfr$. Then branching coefficients $b^{(\mu)}_{\nu}$ for the reduction
$L^{(\mu)}_{\gf}\downarrow \af$ coincide with weight multiplicities $m^{(\tilde \mu)}_{\nu}$ in
$\sfr$-representations provided certain technical condition holds \cite{2011arXiv1111.6787L}.
Highest weight $\tilde\mu$ of $\sfr$-representation is calculated from Dynkin labels of highest
weight $\mu$ of $\gf$-module.

In next section \ref{sec:splints-root-systems} we review the classification of splints for simple
root systems and results for branching coefficients. Then we move to the study of special embeddings
which is the main subject of the present paper. We consider special embeddings of Lie subalgebras
into Lie algebra. In this case root system of the subalgebra is not contained in the root system of
the algebra. So the original motivation for splints is not applicable. But we can consider the
projection of root system of algebra on the root space of the subalgebra. Such a projection is not a
root system anymore, but it satisfies milder conditions (Section \ref{sec:spec-embedd-proj}). It is
possible to classify most projections using Dynkin diagrams augmented with multiplicities. We then
define splint for such 'weak' root systems and state the conditions of its appearance and
implications for the calculation of branching coefficients (Section \ref{sec:splints-spec-embedd}).

@@

In conclusion \ref{sec:conclusion} we discuss the cases when the projection of the root system does
not fall into above mentioned classification.

\section{Splints of root systems and regular subalgebras}
\label{sec:splints-root-systems}
The notion of splint was introduced in the paper \cite{richter2008splints} and classification of
splints for simple root systems was obtained there.

Embedding $\phi$ of a root system $\Delta_1$ into a root system
$\Delta$ is a bijective map of roots of $\Delta_{1}$ to a (proper)
subset of $\Delta$ that commutes with vector composition law in
$\Delta_{1}$ and $\Delta$.
\begin{equation*}
\phi:\Delta_1 \longrightarrow \Delta
\end{equation*}
\begin{equation*}
\phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,
\,\,\, \alpha,\beta \in \Delta_1
\end{equation*}

Note that the image $Im(\phi)$ must not inherit the root system
properties except the addition rules equivalent to the addition
rules in $\Delta_{1}$ (for pre-images). Two embeddings $\phi_1$ and $\phi_2$  
can splinter $\Delta$  when the latter can be presented 
as a disjoint union of images $Im(\phi_1)$ and $Im(\phi_2)$.

The classification is shown in table

Consider a simple Lie algebra $\mathfrak{g}$ and its regular subalgebra $%
\mathfrak{a}\hookrightarrow \mathfrak{g}$ such that $\mathfrak{a}$
is a
reductive subalgebra $\mathfrak{a \subset g}$ with correlated root spaces: $%
\mathfrak{h}_{\mathfrak{a}}^{\ast }\subset \mathfrak{h}_{\mathfrak{g }%
}^{\ast }$. Let $\mathfrak{a}^{\mathfrak{s}}$ be a semisimple summand of
$\mathfrak{a}$,
this means that $\mathfrak{a}=\mathfrak{a}^{\mathfrak{s}} \oplus \mathfrak{u}(1)\oplus %
\mathfrak{u}(1)\oplus \dots$. We shall consider $\mathfrak{a}^{\mathfrak{s}}$
to be a proper regular subalgebra and $\mathfrak{a}$ to be the
maximal subalgebra with $\mathfrak{a}^{\mathfrak{s}}$ fixed that is the rank
$r$ of $\frak{a}$ is equal to that of $\mathfrak{g}$.

Computation of branching coefficients relies on roots $\Delta_{\gf}\setminus \Delta_{\af}$
\cite{2010arXiv1007.0318L} so splint  is naturally connected branching coefficients for regular
subalgebra.

Remember that only three types of splints are
injective and thus are naturally connected with branching. Below we
reproduce the part of the splints table from \cite{richter2008splints} corresponding to
injective splints:
\begin{equation}
\label{eq:1}
\begin{array}{cc||c|c}
\hbox{type} & \hspace{0.25in}\Delta \hspace{0.25in} & \hspace{0.25in}\Delta
_{\frak{a}}\hspace{0.25in} & \hspace{0.25in}\Delta _{\sfr}\hspace{0.25in}
\\ \hline\hline
\hbox{(i)} & G_{2} & A_{2} & A_{2} \\
& F_{4} & D_{4} & D_{4} \\ \hline
\hbox{(ii)} & B_{r}(r\geq 2) & D_{r} & \oplus ^{r}A_{1} \\
(*)& C_{r}(r\geq 3) & \oplus ^{r}A_{1} &  D_{r} \\ \hline
\hbox{(iii)} & A_{r}(r\geq 2) & A_{r-1}\oplus u\left( 1\right)  & \oplus
^{r}A_{1} \\
& B_{2} & A_{1}\oplus u\left( 1\right)  & A_{2}
\end{array}
\end{equation}

@@

\begin{Prop}
\begin{equation}
\frac{e^{\rho _{\gf}}}{\prod_{\beta \in \Delta _{\sfr%
}^{+}}(1-e^{-\beta })}\left( \Psi ^{\widetilde{\mu }+\rho _{\sfr%
}}\right) =\sum_{\widetilde{\nu }\in \mathcal{N}_{\sfr}^{\widetilde{\mu }%
}}M_{\left( \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}e^{\left(
\mu -\phi \left( \widetilde{\mu }-\widetilde{\nu }\right) \right)
}=\sum_{\nu \in P_{\frak{a}}^{++}}b_{\nu }^{(\mu )}e^{\nu }.
\label{singular main-4}
\end{equation}

Actually:
\begin{equation}
\frac{e^{\rho _{\frak{a}}-\rho _{\gf}}}{\prod_{\beta \in \Delta _{\sfr%
}^{+}}(1-e^{-\beta })}\left( \Phi ^{\mu }_{\gf}\right) = \frac{e^{\rho _{\frak{a}}-\rho _{\gf}}}{\prod_{\beta \in \Delta _{\sfr%
}^{+}}(1-e^{-\beta })}\left( e^{\mu +\rho _{\gf}}\phi\left( e^{-\widetilde{\mu }}\Psi^{\widetilde{\mu }}_{\sfr}\right)\right) = 
\end{equation}

\begin{equation}
= e^{\rho _{\frak{a}}+\mu-\phi\left(\widetilde{\mu }\right)}\phi\left(ch L_s^{\widetilde{\mu }}\right) = e^{\rho _{\frak{a}}+\mu-\phi\left(\widetilde{\mu }\right)}\sum_{\widetilde{\nu }\in \mathcal{N}_{\sfr}^{\widetilde{\mu }%
}}M_{\left( \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}e^{\left(
\mu -\phi \left( \widetilde{\mu }-\widetilde{\nu }\right) \right)} =\sum_{\nu \in P_{\frak{a}}^{++}}b_{\nu }^{(\mu )}\Phi_a^{\left(\nu\right)}.
\end{equation}

Any weight with nonzero multiplicity in the r. h. s. is equal to one of the
highest weights in the decomposition. The multiplicity $M_{\left( \sfr%
\right) \widetilde{\nu }}^{\widetilde{\mu }}$ of the weight  $\widetilde{\nu
}\in \mathcal{N}_{\sfr}^{\widetilde{\mu }}$ defines the branching
coefficient $b_{\nu }^{(\mu )}$ for the highest weight $\nu =\left( \mu
-\phi \left( \widetilde{\mu }-\widetilde{\nu }\right) \right) $:
\[
b_{\left( \mu -\phi \left( \widetilde{\mu }-\widetilde{\nu }\right) \right)
}^{(\mu )}=M_{\left( \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}.
\]
\end{Prop}
%% Rewrite this property informally

@@

The study of special embeddings was traces back to fundamental papers by Eugene Dynkin
\cite{dynkin1952semisimple,dynkin1952maximal}, he called such subalgebras ``S-subalgebras'' to
distinguish from regular or ``R-subalgebras'' that have root systems obtained by dropping some roots
of the algebra root system. While regular subalgebras are easy to construct by dropping nodes from
extended Dynkin diagram of the algebra, the case of special subalgebras is more difficult. Complete
classification for exceptional Lie algebras was obtained recently in \cite{minchenko2006semisimple}.
The algorithm of construction of special subalgebras is available in GAP package but still requires
manual intervention \cite{de2011constructing}.

@@

To construct an embedding $\af\to\gf$ consider some representation $L^{\nu}_{\af}$ as a subspace of
Lie algebra $\gf$. Then one needs to check that the generators of $\af$ then can be presented as
linear combinations of generators of $\gf$. We can identify Cartan subalgebra $\hfa\subset \af$ with
dual space $\hfa^{*}$ using Killing form. Then the root system $\Delta_{\gf}$ of $\gf$ can be
projected to $\hfa^{*}$ using the expression of $\hfa$-generators through $\hfg$-generators.

@@

From classical papers \cite{dynkin1952semisimple,dynkin1952maximal} it is well known that the
projection of $\gf$ root system $\Delta_{\gf}$ to $\hfa^{*}$ is a weight system of some
finite-dimensional but not irreducible representation of $\af$.  We denote such a projection by
\begin{equation}
  \label{eq:2}
  \Delta'=\pia\left(\Delta_{\gf}\right)
\end{equation}

The system $\Delta'$ can be a root system but it can be not reduced and can contain some vectors
with multiplicities greater then one.

%Теоремы из Дынкина и их очевидные следствия:

Several theorems from \cite{dynkin1952semisimple} are useful to us.

\begin{theorem}\label{dyn0}
  If representation $\phi$ of algebra $\gf$ induces representation $\tilde\phi$ on subalgebra $\af$ then
  the weight system of representation $\tilde\phi$ can be obtained from representation $\phi$ by
  orthogonal projection of the weight space of algebra $\gf$ on the weight space of subalgebra $\af$

  \cite{dynkin1952maximal}. %(Требует проверки, так как сформулировано в туманных и устаревших терминах в обоих источниках: см. стр.374 в d1 в доказательстве теоремы 3.2 или стр.119 в d2 - теорема 0.11)
\end{theorem}

Applying this theorem to the adjoint representation of $\gf$ we see that the roots of $\af$ are
contained in the projection $\Delta'$ of root system $\Delta_{\gf}$.

Moreover, the following theorems hold. 
\begin{theorem}\label{dyn1}
  Every special subalgebra $\af$ of semisimple Lie algebra $\gf$ is integer, i.e. projections of
  roots of $\gf$ to $\hfa^{*}$ are linear combinations of simple roots of $\af$ with integer
  coefficients \cite{dynkin1952semisimple}
\end{theorem}

\begin{theorem}\label{dyn2}
  If $\af$  is semisimple subalgebra of semisimple Lie algebra $\gf$ and generator $e_{\alpha}$ corresponding to
  a root $\alpha$ of subalgebra $\af$ is presented as a linear combination $e_{\alpha}=\sum_{\beta}
  e_{\beta}$ of $\gf$-generators
  corresponding to roots $\beta$ of $\gf$, then the roots $\beta$ are projected into $\alpha$,
  $\pia(\beta)=\alpha$. 
  \cite{dynkin1972semisimple,dynkin1952semisimple}. 
\end{theorem}

@@

As an example we present a projection $\Delta'$ of root system $D_{4}$ to root system of special
subalgebra $A_{2}$ (Fig. \ref{fig:d4-a2_splint}).

\begin{figure}[h!bt]
  \noindent\centering{
   \includegraphics[width=80mm]{Special-A2-D4-all-2}
  }
  \caption{Projection of $D_{4}$ ($so(8)$)-root system onto root space of special subalgebra $A_{2}$
    ($su(3)$). Note that this roots system is $G_{2}$ but with non-trivial multiplicities. }
 \label{fig:d4-a2_splint}
\end{figure}

@@

Since roots are weights of the adjoint representation, such a projection produces the weight diagram
of $\af$-representation that contains adjoint representation of $\af$. After the subtraction of
$\af$-root system $\Delta_{\af}$ from $\Delta'$ we obtain weight diagram of representation called
``characteristic representation of subalgebra $\af$'' by Dynkin in the seminal paper
\cite{dynkin1952semisimple}.

We want to classify all the cases when branching coefficients coincide with weight multiplicities of
some other algebra representations. Such a coincidence is possible when the projection $\Delta'$ of
$\gf$ root system admits a splint $\Delta'=\varphi_{\af}(\Delta_{\af})\cup
\varphi_{\sfr}(\Delta_{\sfr})$, where $\varphi_{\af}$ and $\varphi_{\sfr}$ are the embeddings of
corresponding root systems. The embedding $\varphi_{\af}$ is metric and trivial. Moreover, the
projection of singular element $\pia\left(\Psi^{\mu}\right)$ should admit a decomposition into
linear combination of singular elements of irreducible representations of $\sfr$:
\begin{equation}
  \label{eq:4}
  \pia\left(\Psi^{\mu}_{\gf}\right)=\sum_{\nu} \Psi^{\nu}_{\sfr}.
\end{equation}
First we classify all the splints
of the projection of root system into union of simple root systems for special embeddings, and then
discuss the decomposition of singular elements.

Projection of the root system $\Delta'$ coincides with the projection of weight diagram of the
adjoint representation of algebra $\gf$ with the exception of zero weight. Adjoint representation of
$\gf$ contains adjoint representation of $\af$ and can be decomposed as
\begin{equation}
  \label{eq:3}
  \mathrm{ad}_{\gf}=\mathrm{ad}_{\af}\oplus M_{\af}^{\chi},
\end{equation}
where $M^{\chi}_{\af}$ is not necessary irreducible representation of $\af$ called
``characteristic'' in \cite{dynkin1952semisimple}.

Weight diagram of $M^{\chi}_{\af}$ coincides with root system $\Delta_{\sfr}$ with the exception of
zero weight. So we need to find all the representations for all simple Lie algebras $\af$, such that
their weight diagrams are root systems after the exclusion of zero weight. In order to do so we must
note, that all the weights of $M^{\chi}_{\af}$ should have length not less than length of shortest
root of $\af$, since $\af$ is an integer subalgebra of $\gf$ \ref{dyn1}. Moreover, $M^{\chi}_{\af}$
should contain weights of at most two different lengths.

>>>>

\usepackage{amsfonts}
\usepackage{pb-diagram}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\usepackage{amsmath,amssymb,amsthm}
\usepackage{multicol}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}

@@

Note that the image $Im(\phi)$ is not required to inherit the root system properties except the
addition rules equivalent to the addition rules in $\Delta_{1}$ (for pre-images). If an embedding
$\phi$ preserves the angles between the roots it is called ``metric''. Two embeddings $\phi_1$ and
$\phi_2$ can splinter $\Delta$ when the latter can be presented as a disjoint union of images
$Im(\phi_1)$ and $Im(\phi_2)$.

Why one would consider non angle-preserving maps of root systems? Additive properties of root system
determine the structure of Verma module:
    \begin{equation*}
      M^{\mu}=U(\gf)\underset{U(\bff_{+})}{\otimes} D^{\mu}(\bff_{+})=\{(E^{-\alpha_{1}})^{n_{1}}\dots (E^{-\alpha_{s}})^{n_{s}} \left|v_{\mu}\right>\}_{\alpha_{i}\in\Delta^{+}}^{n_{i}=0,1,\dots}
    \end{equation*}
Here $\bff_{+}$ is the Borel subalgebra of $\gf$, $D^{\mu}$ its one-dimensional representation,
$\left|v_{\mu}\right>$ is the highest weight vector and $E^{-\alpha_{j}}$ are lowering operators
that are in correspondence with positive roots $\alpha_{j}\in \Delta^{+}$. 
Weyl character formula expresses character of irreducible representation as a combination of
characters of Verma modules: 
\begin{equation*}
  \mathrm{ch} L^{\mu}=\frac{\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}}{\sum_{w\in W}\varepsilon(w) e^{w\rho-\rho}}=\sum_{w\in W} \varepsilon(w)\; \mathrm{ch} M^{w(\mu+\rho)-\rho}
\end{equation*}
Here the sum is over elements of the Weyl group and their actions $w\triangleright \mu
=w(\mu+\rho)-\rho$ depend on angles between roots. But we can rewrite Weyl character formula
representing Weyl group elements as the products of reflections $s_{\alpha}, \alpha \in S$ in
hyperplanes orthogonal to simple roots: $w=s_{\alpha_{1}}\cdot s_{\alpha_{2}}\dots$. Reflections act
on root system by permutations, so the composition of such an action with embedding $\phi$ is easily
obtained. The orbit of Weyl groups action $w\triangleright \mu, w\in W$, can be constructed by
subtractions of roots from the highest weight $\mu$. Let's denote Dynkin labels of $\mu$ by
$(\mu_{1},\dots \mu_{r})$. Then $\mu-\mu_{1}\alpha_{1}, \mu-\mu_{2}\alpha_{2},\dots,
\mu-\mu_{r}\alpha_{r}, \alpha_{i}\in S$, are the points of the orbit that can be obtained by
elementary reflections $s_{\alpha_{i}}, \alpha_{i}\in S$. Next set of points that are obtained by
two consecutive reflections $s_{\alpha_{i}}\cdot s_{\alpha_{j}}\triangleright\mu$ are obtained as
the subtraction $\mu-\mu_{i}\alpha_{i}-\mu_{j} (s_{\alpha_{i}}\alpha_{j})$. Continuing this way we
construct the image of singular element $\Psi=\sum_{w\in W} \varepsilon(w) e^{w(\mu+\rho)-\rho}$
after the embedding $\phi$ \cite{2011arXiv1111.6787L}. The multiplicities in the character of Verma
module are unchanged by the embedding since they are determined by additive properties of the roots.
So we see that weight multiplicities of irreducible modules are also preserved by the embedding.

@@

Assume that root system $\Delta_{\gf}$ splinters as $\Delta_{\gf}=\Delta_{\af} \cup
\phi(\Delta_{\sfr})$, where $\phi$ is an embedding of root system $\Delta_{\sfr}$ of some semisimple
Lie algebra $\sfr$. Then branching coefficients $b^{(\mu)}_{\nu}$ for the reduction
$L^{(\mu)}_{\gf\downarrow \af}$ coincide with weight multiplicities $m^{(\tilde \mu)}_{\nu}$ in
$\sfr$-representations provided certain technical condition holds \cite{2011arXiv1111.6787L}.
Highest weight $\tilde\mu$ of $\sfr$-representation is calculated from Dynkin labels of highest
weight $\mu$ of $\gf$-module.

In the next section \ref{sec:splints-root-systems} we review the classification of splints for simple
root systems and results for branching coefficients. Then we move to the study of special embeddings
which is the main subject of the present paper. We consider special embeddings of Lie subalgebras
into a Lie algebra. In this case root system of the subalgebra is not contained in the root system of
the algebra. So the original motivation for splints is not applicable. But we can consider the
projection of root system of the algebra on the root space of the subalgebra. Such a projection is not a
root system anymore, but it satisfies milder conditions (Section \ref{sec:spec-embedd-proj}). It is
possible to classify most projections using Dynkin diagrams augmented with multiplicities. We then
define splint for such 'weak' root systems and state the conditions of its appearance and
implications for the calculation of branching coefficients (Section \ref{sec:splints-spec-embedd}).

@@

In conclusion \ref{sec:conclusion} we discuss the cases when the projection of the root system does
not fall into the classification mentioned above.

\section{Splints of root systems and regular subalgebras}
\label{sec:splints-root-systems}
The notion and classification of
splints for simple root systems were introduced in the paper \cite{richter2008splints}.

%% Embedding $\phi$ of a root system $\Delta_1$ into a root system
%% $\Delta$ is a bijective map of roots of $\Delta_{1}$ to a (proper)
%% subset of $\Delta$ that commutes with vector composition law in
%% $\Delta_{1}$ and $\Delta$.
%% \begin{equation*}
%% \phi:\Delta_1 \longrightarrow \Delta
%% \end{equation*}
%% \begin{equation*}
%% \phi \circ (\alpha + \beta) =\phi \circ \alpha + \phi \circ \beta,
%% \,\,\, \alpha,\beta \in \Delta_1
%% \end{equation*}
%% 
%% Note that the image $Im(\phi)$ must not inherit the root system
%% properties except the addition rules equivalent to the addition
%% rules in $\Delta_{1}$ (for pre-images). Two embeddings $\phi_1$ and $\phi_2$  
%% can splinter $\Delta$  when the latter can be presented 
%% as a disjoint union of images $Im(\phi_1)$ and $Im(\phi_2)$.   
%% 
%% The classification is shown in table
%% 
Consider a simple Lie algebra $\mathfrak{g}$ and its regular subalgebra $%
\mathfrak{a}\hookrightarrow \mathfrak{g}$ such that $\mathfrak{a}$
is a
reductive subalgebra $\mathfrak{a \subset g}$ with correlated root spaces: $%
\mathfrak{h}_{\mathfrak{a}}^{\ast }\subset \mathfrak{h}_{\mathfrak{g }%
}^{\ast }$. Let $\mathfrak{a}^{\mathfrak{s}}$ be a semisimple summand of
$\mathfrak{a}$,
this means that $\mathfrak{a}=\mathfrak{a}^{\mathfrak{s}} \oplus \mathfrak{u}(1)\oplus %
\mathfrak{u}(1)\oplus \dots$. We shall consider $\mathfrak{a}^{\mathfrak{s}}$
to be a proper regular subalgebra and $\mathfrak{a}$ to be the
maximal subalgebra with $\mathfrak{a}^{\mathfrak{s}}$ fixed that is the rank
$r$ of $\frak{a}$ is equal to that of $\mathfrak{g}$.

Computation of branching coefficients relies on roots $\Delta_{\gf}\setminus \Delta_{\af}$
\cite{2010arXiv1007.0318L} so splint  is naturally connected to branching coefficients for regular
subalgebra.

Only three types of splints are
injective and thus are naturally connected to branching \cite{richter2008splints}:
\begin{equation}
\label{eq:1}
\begin{array}{cc||c|c}
\hbox{type} & \hspace{0.25in}\Delta \hspace{0.25in} & \hspace{0.25in}\Delta
_{\frak{a}}\hspace{0.25in} & \hspace{0.25in}\Delta _{\sfr}\hspace{0.25in}
\\ \hline\hline
\hbox{(i)} & G_{2} & A_{2} & A_{2} \\
& F_{4} & D_{4} & D_{4} \\ \hline
\hbox{(ii)} & B_{r}(r\geq 2) & D_{r} & \oplus ^{r}A_{1} \\
(*)& C_{r}(r\geq 3) & \oplus ^{r}A_{1} &  D_{r} \\ \hline
\hbox{(iii)} & A_{r}(r\geq 2) & A_{r-1}\oplus u\left( 1\right)  & \oplus
^{r}A_{1} \\
& B_{2} & A_{1}\oplus u\left( 1\right)  & A_{2}
\end{array}
\end{equation}

@@

\begin{Prop}
There is a decomposition of a singular element of the algebra $\gf$ into singular elements of the subalgebra $\sfr$:
\begin{equation}
 \Phi ^{\mu }_{\gf} = \sum_{w\in W_{\frak{a}}}\epsilon\left( w\right)w\left( e^{\mu +\rho _{\gf}}\phi\left( e^{-\widetilde{\mu }}\Psi^{\widetilde{\mu }}_{\sfr}\right)\right),
\end{equation}
where $\epsilon\left( w\right)$ is a determinant of the element $w$ of the Weyl group $W_{\af}$ of
the subalgebra $\af$. The action of Weyl group on the weights is extended to the algebra of formal
exponents by the rule $w\left(e^{\nu}\right)=e^{w\nu}, w\in W$. Hence, the multiplicity $M_{\left(
      \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}$ of the weight $\widetilde{\nu }$ from the
  weight diagram of the algebra $\sfr$ with the highest weight $\tilde\mu$ defines the branching
  coefficient $b_{\nu }^{(\mu )}$ for the highest weight $\nu =\left( \mu -\phi \left(
      \widetilde{\mu }-\widetilde{\nu }\right) \right) $:
\begin{equation}
b_{\left( \mu -\phi \left( \widetilde{\mu }-\widetilde{\nu }\right) \right)
}^{(\mu )}=M_{\left( \sfr\right) \widetilde{\nu }}^{\widetilde{\mu }}. 
\label{bran1}
\end{equation}
\end{Prop} 
%% Rewrite this property informally

@@

The study of specialy embedded subalgebras traces back to fundamental papers by Eugene Dynkin
\cite{dynkin1952semisimple,dynkin1952maximal} where he called such subalgebras ``S-subalgebras'' to
distinguish from regular or ``R-subalgebras'' that have root systems obtained by dropping some roots
of the algebra root system. Thus regular subalgebras are easy to construct by dropping nodes from
extended Dynkin diagram of the algebra, while the case of special subalgebras is more difficult. Complete
classification for exceptional Lie algebras was obtained recently in \cite{minchenko2006semisimple}.
The algorithm of the special subalgebras construction is available in GAP package but still requires
manual intervention \cite{de2011constructing}.

@@

To construct an embedding $\af\to\gf$ consider some representation $L^{\nu}_{\af}$ as a subspace of
Lie algebra $\gf$. Then one needs to check that the generators of $\af$ can be presented as
linear combinations of generators of $\gf$. We can identify Cartan subalgebra $\hfa\subset \af$ with
dual space $\hfa^{*}$ using Killing form. Then the root system $\Delta_{\gf}$ of $\gf$ can be
projected to $\hfa^{*}$ using the expression of $\hfa$-generators through $\hfg$-generators.

@@

This projection is described in the classical papers \cite{dynkin1952semisimple,dynkin1952maximal} by the following theorem:

\begin{theorem}\label{dyn0}
  If a representation $L^{\mu}_{\gf}$ of the algebra $\gf$ induces a representation
  $L^{\tilde\mu}_{\af}$ on the subalgebra $\af$ then the weight system of $L^{\tilde\mu}_{\af}$ can
  be obtained from $L^{\mu}_{\gf}$ by an orthogonal projection of $\hfg^{*}$ on $\hfa^{*}$.

  \cite{dynkin1952maximal}. 
\end{theorem}

In relation to the adjoint representation of $\gf$ this theorem means that the projection of the root system $\Delta_{\gf}$ to $\hfa^{*}$ is a weight system of some
finite-dimensional but not irreducible representation of $\af$. Moreover this reducible representation contains the adjoint representation of $\af$. And if we denote such a projection by
\begin{equation}
  \label{eq:2}
  \Delta'=\pia\left(\Delta_{\gf}\right),
\end{equation}
then the roots of $\af$ will be contained in $\Delta'$. Thus the system $\Delta'$ can be a root
system but it might not be reduced and might contain some vectors with multiplicities greater than
one.

The following theorems \cite{dynkin1952semisimple} also concern the properties of $\Delta'$.

\begin{theorem}\label{dyn1}
  Every special subalgebra $\af$ of a semisimple Lie algebra $\gf$ is integer, i.e. the projections of
  roots of $\gf$ to $\hfa^{*}$ are linear combinations of simple roots of $\af$ with integer
  coefficients \cite{dynkin1952semisimple}
\end{theorem}

\begin{theorem}\label{dyn2}
  If $\af$  is a semisimple subalgebra of a semisimple Lie algebra $\gf$ and the generator $e_{\alpha}$ corresponding to
  the root $\alpha$ of $\af$ is presented as a linear combination $e_{\alpha}=\sum_{\beta}
  e_{\beta}$ of $\gf$-generators
  corresponding to the roots $\beta$ of $\gf$, then the roots $\beta$ are projected into $\alpha$,
  $\pia(\beta)=\alpha$. 
  \cite{dynkin1972semisimple,dynkin1952semisimple}. 
\end{theorem}

@@

As an example we present a projection $\Delta'$ of the root system of $D_{4}$ to the root system of the special
subalgebra $A_{2}$ (Fig. \ref{fig:d4-a2_splint}).

\begin{figure}[h!bt]
  \noindent\centering{
   \includegraphics[width=80mm]{Special-A2-D4-all-2}
  }
  \caption{Projection of $D_{4}$ ($so(8)$)-root system onto the root space of the special subalgebra $A_{2}$
    ($su(3)$). Note that this roots system is $G_{2}$ but with non-trivial multiplicities. }
 \label{fig:d4-a2_splint}
\end{figure}

@@

%Since roots are weights of the adjoint representation, such a projection produces the weight diagram
%of $\af$-representation that contains adjoint representation of $\af$.
%After the subtraction of
%$\af$-root system $\Delta_{\af}$ from $\Delta'$ we obtain the weight diagram of representation called
%``characteristic representation of subalgebra $\af$'' by Dynkin in the seminal paper
%\cite{dynkin1952semisimple}.

We want to classify all the cases when branching coefficients coincide with weight multiplicities of
some other algebra representations. Such a coincidence is possible when the projection $\Delta'$ of
$\gf$ root system admits a splint $\Delta'=\varphi_{\af}(\Delta_{\af})\cup
\varphi_{\sfr}(\Delta_{\sfr})$, where $\varphi_{\af}$ and $\varphi_{\sfr}$ are the embeddings of
corresponding root systems. The embedding $\varphi_{\af}$ is metric and trivial. Moreover, the
projection of singular element $\pia\left(\Psi^{\mu}_{\gf}\right)$ should admit a decomposition into
linear combination of singular elements of irreducible representations of $\sfr$:
\begin{equation}
  \label{eq:4}
  \pia\left(\Psi^{\mu}_{\gf}\right)=\sum_{\nu} \varkappa_{\nu}e^{\nu}\Psi^{\tilde\mu}_{\sfr},
\end{equation}
with integer coefficients $\varkappa_{\nu}$, where the sum is over the set of weights $\nu$ that
will be determined later. First we classify all the splints of the projection of a root system into
a union of simple root systems for special embeddings, and then discuss the decomposition of singular
elements.

Projection of the $\gf$ root system $\Delta'$ coincides with the projection of the weight diagram of the
adjoint representation of algebra $\gf$ with the exception of zero weight. The adjoint representation of
$\gf$ contains adjoint representation of $\af$ and can be decomposed as
\begin{equation}
  \label{eq:3}
  \mathrm{ad}_{\gf}=\mathrm{ad}_{\af}\oplus M_{\af}^{\chi},
\end{equation}
where $M^{\chi}_{\af}$ is not necessary irreducible representation of $\af$ called
``characteristic'' in \cite{dynkin1952semisimple}.

The weight diagram of $M^{\chi}_{\af}$ coincides with the root system $\Delta_{\sfr}$ with the exception of
zero weight. So we need to find all the representations for all simple Lie algebras $\af$, such that
their weight diagrams are root systems after the exclusion of zero weight. In order to do so we must
note, that all the weights of $M^{\chi}_{\af}$ should have length not less than length of shortest
root of $\af$, since $\af$ is an integer subalgebra of $\gf$ \ref{dyn1}. Moreover, $M^{\chi}_{\af}$
should contain weights of at most two different lengths.

If the root system $\Delta_{\gf}$ contains roots $\alpha: \alpha\perp \beta,\; \forall \beta\in
\Delta_{\af},$ orthogonal to the root system of the subalgebra, one cannot proceed with the decomposition
of singular element \eqref{eq:4}. The elements $e^{\nu}$ of $\pia\left(\Psi^{\mu}_{\gf}\right)$
should be augmented with the dimensions of representations of algebra $\afb$ spanned by generators
corresponding to orthogonal roots \cite{2010arXiv1007.0318L}. Then there will be non-trivial
multiplicities in the right hand side of \eqref{eq:4} and no coincidence of branching coefficients
with weight multiplicities of $\sfr$-representations. One can write more cumbersome relation between
multiplicities and branching coefficients, but it is out of scope for the present paper.
